---
title: "Summary stats on mexican subsidy, landings, and effort data"
subtitle: "Disclaimer: I'm still getting to know and clean the data"
author: "Juan Carlos Villaseñor-Derbez"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

suppressPackageStartupMessages({
  library(ggridges)
  library(here)
  library(tidyverse)
})

eu_panel <- readRDS(here("data", "economic_unit_subsidy_panel.rds"))
theme_set(startR::ggtheme_plot())
```

# Subsidy data

There are `r length(unique(eu_panel$rnpa))` "Economic units" (cooperatives, individuals, companies) that received a fuel subsidy at least once between 2011 and 2019.

```{r}

mean_sd <- function(x){
  mean <- mean(x, na.rm = T)
  sd <- sd(x, na.rm = T)
  
  c("y" = mean,
    "ymin" = mean - sd,
    "ymax" = mean + sd)
}

eu_panel %>% 
  filter(subsidy_amount > 0) %>% 
  ggplot(aes(x = year, y = subsidy_amount, group = year)) +
  stat_summary(geom = "errorbar", fun.data = mean_sd, width = 0, size = 1) +
  stat_summary(geom = "point", fun = mean, color = "black", fill = "steelblue", shape = 21, size = 3) +
  facet_wrap(~fuel_type, ncol = 1, scale = "free_y") +
  labs(x = "Year", y = "Awared subsidy (Mean +- SD)", title = "Mean annual subsidy amount, by fuel type")
```


```{r}
non_zero_subsidy <- eu_panel %>% 
  filter(subsidy_amount > 0) %>% 
  group_by(rnpa, fishing_type) %>% 
  summarize(n = n(),
            mean_s = mean(subsidy_amount),
            sd_s = sd(subsidy_amount))

eu_panel %>% 
  left_join(non_zero_subsidy, by = c("rnpa", "fishing_type")) %>% 
  filter(subsidy_amount > 0) %>% 
  mutate(norm_sub = subsidy_amount - mean_s) %>% 
  ggplot(aes(x = year, y = norm_sub, group = year)) +
  geom_jitter(height = 0, width = 0.25, alpha = 0.2, size = 0.1) +
  facet_wrap(~fuel_type, ncol = 1) +
  scale_y_continuous(limits = c(-1, 1) * 1e6) +
  labs(x = "Year", y = expression(Subsidy[it]-mean[i]), title = "Vessel-level deviations from the mean")
  
  
```


```{r}
eu_panel %>% 
  filter(subsidy_amount > 0) %>% 
  count(rnpa, fishing_type) %>% 
  count(n) %>% 
  ungroup() %>% 
  mutate(pct = cumsum(nn) / sum(nn)) %>% 
  ggplot(aes(x = n, y = nn)) +
  geom_col() +
  labs(x = "Number of times subsidied",
       y = "Number of vessels",
       title = "How many times was each vessel subsidised?")
```


```{r}
eu_panel %>% 
  filter(subsidy_amount > 0) %>% 
  count(rnpa, fishing_type) %>% 
  count(n) %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  mutate(pct = cumsum(nn) / sum(nn)) %>% 
  ggplot(aes(x = n, y = pct)) +
  geom_line() +
  geom_point(fill = "steelblue", size = 3, shape = 21) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = 0:11, labels = c(0:11)) +
  labs(x = "Number of times subsidised",
       y = "Number of vessels",
       title = "Y vessels received a subsidy at least X times (for the subsidiesd ones)") 
```


# Landings data

```{r}
landings <- readRDS(here("data", "landings_clean.rds"))
```

```{r}
a <- eu_panel %>% 
  left_join(landings, by = c("year" = "ano_corte", "rnpa" = "rnpa_unidad_economica"))
```


A quick and dirty match says there are `r sum(!is.na(a$landings))` observations with a match and `r sum(is.na(a$landings))` observations without a match. That is a match of `r (1 - sum(is.na(a$landings)) / length(a$landings)) * 100`\%.

```{r}

hs <- function(x){
  log(x + sqrt((x ^ 2 + 1)))
}

a %>% 
  drop_na(landings, subsidy_amount) %>% 
  ggplot(aes(x = hs(subsidy_amount), y = log(landings))) +
  geom_point(alpha = 0.2) +
  labs(x = "Subsidy (inverse-hyperbolic-sine transformed)", y = "Log (landings)",
       title = "Years 2011 - 2018")
```


# Some regressions

$$log_{10}(L_{i,t}) = \alpha + \beta_1 \mathrm{ihs}(S_{i,t}) + \gamma_tY_t + \phi_iRNPA + \epsilon_{it}$$
```{r}

list(fixest::feols(log10(landings) ~ hs(subsidy_amount) + fuel_type | as.factor(year) + as.factor(rnpa), data = a))%>% 
  modelsummary::modelsummary(gof_omit = "Pseudo|Adj|With|IC|Log", stars = TRUE, statistic_vertical = FALSE, output = "latex")

```


# Effort data

```{r}
effort <- readRDS(here("data", "vessel_anual_effort.rds"))

effort_ais <- readRDS(here("data", "vessel_anual_effort_AIS.rds"))
```

There are `r length(unique(effort$effort$ssvid))` vessels in the VMS dataset (1 ping per hour). There are `r length(unique(effort_ais$ssvid))` vessels in the AIS dataset (Many more pings).


```{r}
ggplot(effort, aes(x = year, y = hours)) +
  stat_summary(geom = "errorbar", fun.data = mean_sd, width = 0, size = 1) +
  stat_summary(geom = "point", fun = mean, color = "black", fill = "steelblue", shape = 21, size = 3) +
  labs(x = "Year", y = "Fishing hours (Mean +- SD)", title = "Mean annual fishing hours (VMS, 2014 - 2016)")
```

```{r}
ggplot(effort_ais, aes(x = year, y = fishing_hours)) +
  stat_summary(geom = "errorbar", fun.data = mean_sd, width = 0, size = 1) +
  stat_summary(geom = "point", fun = mean, color = "black", fill = "steelblue", shape = 21, size = 3) +
  labs(x = "Year", y = "Fishing hours (mean +- SD)", title = "Mean annual fishing hours (VMS, 2014 - 2016)")
```


# Fuel prices

I have not been able to find fuel prices per year, so I'm using crude oil prices from Mexican oil production. That will be fixed evenutally.

```{r}
fuel <- read.csv(here("data", "oil_prices.csv")) %>% 
  pivot_longer(cols = c(Brent_price, WTI_price, MME_price), names_to = "mix", values_to = "price") %>% 
  group_by(year) %>% 
  summarize(price = mean(price, na.rm  = T))
```

```{r}
fuel %>% 
  ggplot(aes(x = year, y = price)) +
  geom_line() +
  geom_point(fill = "steelblue", size = 3, shape = 21) +
  labs(x = "Year", y = "Price (USD / Bbl)")
```


```{r}
effort %>% 
  left_join(fuel, by = "year") %>% 
  ggplot(aes(x = hours, y = price)) +
  geom_point()
```

```{r}
effort_ais %>% 
  left_join(fuel, by = "year") %>% 
  ggplot(aes(x = fishing_hours, y = price)) +
  geom_point() +
  labs(x = "Fishing hours (AIS)", y = "Oil price (USD / BBl)")
```




\clearpage

## Preview of subsidy data

```{r}
eu_panel %>% 
  select(year, rnpa, fishing_type, subsidy_amount, everything()) %>% 
  head(20)
```

\clearpage


We actually have these at the species-level but I'm combining it here. We also have the exvessel price of these landings.

```{r}
head(landings, 20)
```

\clearpage

VMS data

```{r}
head(effort, 20)
```

\clearpage

AIS data

```{r}
head(effort_ais, 20)
```










